# Bootcamp Practices

## Bootcamp Lesson 1

### Practice with
- Variables
- Functions
- Printing to the screen

1. Assign the number 500 to the variable numberButterflies
```{r}
#
```


2. Calculate the cubed root of 26 and assign it to the variable y
```{r}
#
```


3. Print the variable y
```{r}
#
```


## Bootcamp Lesson 2

### Practice with
- Packages
- Accessing specific columns in a data set
- Assigning variables
- Select
- Filtering


1. (Install if you haven't completed lesson 2) and load the tidyverse and palmerpenguins packages. 
```{r}
#
```


2. Assign the bill length column to a variable called penguinBillLength
```{r}
#
```


3. Assign the bill depth column to a variable called penguinBillDepth. 
```{r}
#
```

For Questions 4 - 7, you can choose to either use the pipe %>% or not. Either way is great! 

4. Assign only the columns island, bill_length_mm, and bill_depth_mm to a new variable (name it whatever you'd like!). 
```{r}
#
```

5. Assign only the columns sex and year to the variable sexByYear. 
```{r}
#
```


6. Create a subset of the data that includes all the columns only for male Gentoo penguins with flipper lengths greater than 20 mm. 
```{r}
#
```


7. Create a subset of the data that includes all penguins on Biscoe island with flipper lengths greater than 20 mm. Remove all of the NA values as well. 
```{r}
#
```


Challenge: Using the pipe %>%, string together two commands with only one line of code to create a data set that includes only the columns: species, sex, and flipper length, with only Adelie penguins whose flippers are greater than 180 mm. 
```{r}
#
```



## Bootcamp Lesson 3

### Practice with
* Packages
* Filtering
* Ggplot

### Instructions for plots
Each plot should have axis labels, a title, and use colors to illustrate the data.


1. (Install if you haven't completed lessons 2 or 3 or practice 2) and load the tidyverse and palmerpenguins packages. 
```{r}
#
```


2. Create a subset of the data that includes all the data from the years 2007 and 2008.
```{r}
#
```


3. Create a box plot of the bill lengths by male and female penguins of all species from the years 2007 and 2008. 
```{r}
#
```


4. Create a subset of the data that includes only the female penguins. 
```{r}
#
```


5. Create a box plot of the bill depth by species of only the female penguins. 
```{r}
#
```


6. Create a scatter plot of the relationship between bill length and bill depth for penguins on only Dream island and color by species.
```{r}
#
```
